---
title: "What to Expect From Team Brasil in 'Tour Brasil Na WNBA'"
date: "5/2/2025"
author: "Billy Fryer"
categories: ["WNBA"]
tags: ["WNBA", "Women's Basketball", "International"]
---

This WNBA Preseason, Team Brazil takes on professional opponents of the Chicago Sky (May 2) and the Indiana Fever (May 4) in their "Brazilian Tour in the WNBA". This tour serves as practice for this summer's upcoming FIBA Americup in late June/early July in Santiago, Chile where the Brazilians will be slotted in Group A with the Dominican Republic, Argentina, Canada and El Salvador. Below is the tweet posted by @basquetebrasil, the official X account of the Confederação Brasileira de Basketball (Brazilian Basketball Confederation)


```{r tweet, echo=FALSE, message=FALSE, warning=FALSE}
library(tweetrmd)

tweetrmd::include_tweet("https://x.com/basquetebrasil/status/1918124798581907621")
```

# Missing Pieces

Due to WNBA obligations, the Brazilian National Team will be without several key pieces in these preseason match ups. The team announced Pokey Chatman, Associate General Manager and Assistant Coach for the Seattle Storm, will be their head coach for this Olympic Cycle culminating in the 2028 Los Angeles Olympics.

The team will also be missing several post players from their announced roster for this, the biggest being Kamilla Cardoso-Silva. Standing at 6' 7", Cardoso will be suiting up against her National Team in the first match up as she is a projected starter for the Sky this year. She started 29 of the 32 games she participated in last season, showing good efficiency around the rim shooting 55% from within 5 feet on 6.3 attempts per game.

In Team Brazil's second match up with the Indiana Fever, they will be playing against another Brazilian WBB icon: Damiris Dantas. Dantas is a 6' 3" pick and pop post player with 9 years of WNBA experience including several years in Minnesota with Team USA Head Coach Cheryl Reeve. Dantas has been a staple on the Brazilian National Team since her debut with the Senior Team in 2010. In that time span, she's averaged 12 points per game, including 16.8 points per game in the 2016 Rio Olympics.

Another Brazilian WBB stalwart missing from this preliminary roster is Clarissa dos Santos. Dos Santos last appeared for Team Brazil in 2021 edition of the AmeriCup but has appeared in clubs around the world since, most recently in Portugal for Benfica. The 6 foot forward's leadership will surely be missed by As Canarinhas.

# NCAA Student Athletes

```{r college-table, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(gt)
library(gtExtras)
library(cfbplotR)

college_stats <- read_excel("college-stats.xlsx")

college_stats$Team <- college_stats %>% 
  cfbplotR::clean_school_names(school = .$Team) 

college_stats %>%
  select(-contains("2P")) %>% 
  gt() %>% 
  # Convert to logos
  cfbplotR::gt_fmt_cfb_logo(Team) %>% 
  # Format Percents
  fmt_percent(columns = contains("%"),
              decimals = 1) %>% 
  # Headers, footers
  tab_header(title = "Team Brazil during 2024-25 NCAA Season") %>% 
  gt::tab_footnote("Vizualtion by Billy Fryer ~ Data from CBB Ref") %>% 
  # Theme
  gtExtras::gt_theme_espn()
```

4 players on the tour roster logged minutes in NCAA play this season, including 3 with NCAA Tournament experience. Nascimento led the charge for the Tribe of William and Mary with 16.6 PPG and 31 MPG. She took the Tribe from a 9th seed going into the Coastal Athletic Association Tournament where she averaged almost 20 PPG and took home Tournament MVP. This led the Tribe to the NCAA Tournament where they beat High Point in the First Four matchup before falling to Texas in the Round of 64. 

Iza Nicoletti served more of a pass-first guard with the Fairfield Stags, feeding her posts (nicknamed the Roadrunners) in transition with 5.1 assists per game, 2nd in the MAAC. Fairfield played well the entire season, winning both the regular season and tournament championships before falling to Kansas State in the Round of 64 for the second straight season.

Catarina Ferreira in her first season with Oregon State put up career numbers in all categories. Expect this Tour to propel her to an even better Senior year in Corvallis.

Varejão in her fourth NCAA season (second at Syracuse) also put up career numbers in both points and rebounds. Standing at 6' 4", she should replace some of the production lost by Cardoso, Dantas and Dos Santos in the middle.

# Rising Stars

Brazil has 3 players on their roster entering NCAAA competition next season, all of which should be a good post rotation for As Canarinhas for this tour. Though it may be doubtful that these players make the roster for the AmeriCup, it is very great experience going against former NCAA Champions and South Carolina Gamecocks Aliyah Boston of the Fever and Kamilla Cardoso of the Sky.

Ayla McDowell, committed to the reigning National Runner Up South Carolina Gamecocks, is an 18 year old, 6' 2" power forward 4 or 5 star recruit depending on the source. During the U18 Americas Championship last summer, she averaged 14.8 PPG and 7.7 RPG.

Taissa Queiroz was also a highly sought after recruit, now committed to University of North Carolina. She has made 4 appearances in FIBA Youth Competitions averaging a double-double with 11.1 PPG and 10.4 PPG. Her last appearance was at the FIBA U19 World Cup where her team placed 14th. 

Manuella Alves is a 4 star recuit out of IMG Academy in Florida committed to Illinois. Standing at 6' 3", she's played in 6 FIBA Youth Competitions. In 2023, she helped lead her team to the gold medal in the South American U17 Women's Championship where she was named to the All-Star 5 and averaged 9.6 PPG and 9.8 RPG.

# Other Players

```{r pro-table, echo=FALSE, message=FALSE, warning=FALSE}

pro_stats <- read_excel("pro-stats.xlsx")


pro_stats %>%
  gt() %>% 
  # Headers, footers
  tab_header(title = "Team Brazil during 2024-25 Professional Season") %>% 
  gt::tab_footnote("Vizualtion by Billy Fryer ~ Data from Eurobasket.com") %>% 
  # Theme
  gtExtras::gt_theme_espn()
```

The rest of Team Brazil's roster is comprised of mostly players from the top 3 teams in Brazilian Domestic League. The league is currently about halfway through (Currently in Round 11 of 21) with Sampaio currently at 9-1 including a win over second place SESI.

# Thoughts/Predictions

It's fair to conclude that both the Chicago Sky and the Indiana Fever will have no problem defeating Team Brazil in Baton Rouge and Iowa City respectively. Neither WNBA team expects to use traditional rotations as this is preparation for the upcoming training camp cuts that are due by May 15. Indiana's guard play between Caitlin Clark, Kelsey Mitchell, and DeWanna Bonner (among others) likely will overwhelm the Brazilian Guards. However, I expect the game vs the Chicago Sky to be much closer as the Sky adapt to Tyler Marsh performing head coaching duties for the first time in his career.

In both matchups, Team Brazil facing stronger competitions than they are expected to during the AmeriCup this summer. Team USA traditionally sends a College All Star Team to the event when they compete in it. When last played in 2023, Team Brazil defeated Team USA both times they competed against one another with Team USA's roster featuring current LA Spark Rickea Jackson, Chicago Sky player Angel Reese, UCLA Center Lauren Betts and several others still competing in the NCAA or that have logged WNBA minutes.

