---
title: "intRo 1- Basics"
author: "Billy Fryer"
date: 2021-01-20
categories: ["R"]
tags: ["intRo Tutorials"]
---

Here is a link to the recorded version of this tutorial!

[intRo 1 Tutorial Link](https://www.youtube.com/watch?v=rE4ah0nBouY&list=PLzfY2UeNJip_GBjX8xEcRcc9omZjl1jFU&index=1)

# Single Value Variables

```{r Q1}
a <- 10 # this is an INTEGER variable
b <- "Wolfpack" # this is a CHARACTER variable (aka a string)
a
print(a)
b
print(b)
```

# Vectors

```{r vectors}
vec1 <- c(1,2,4,5) 
# An integer vector - the c() function is very useful!
vec2 <- 1:5 # another integer vector, the : makes a sequence from 1:5
vec3 <- c("NC State","Wolfpack") # this is a character vector
print(vec1)
print(vec2)
print(vec3)

class(vec1) 
# To see what type a variable/vector is, use the class() function

# What if we want to look at just a certain item from a vector?
vec1[3] # print the 3rd item/element from vec1
vec3[1] # print the 1st item/element from vec3
# SINGLE brackets

# Can we mix variable types in a vector? No
mixedvec <- c(1:4,"Wolfpack") 
# Because Wolfpack is character, all are converted to character
print(mixedvec)
class(mixedvec)

#' There is another data type called a List that we will not discuss
#' It CAN hold both numeric and character values in a similar way to 
#' vectors. But we want to do more interesting things, so we'll move on
#' for now.
```

# Multidimensional Objects

```{r mdobjects}
mat <- matrix(1:12, nrow=4, ncol=3) # this is a MATRIX
mat
mat[2,3] # extracts item in 2nd ROW, 3rd COLUMN of the matrix
mat[4,] # extracts the 4th row, ALL columns
mat[,2] # extracts ALL rows, the 2nd column

# this is a data.frame
df <- data.frame(Name = c("Tom Brady", "Jackie Robinson", "Kobe Bryant"), 
                 Number = c(12, 42, 24))
df
df[2,2] # You can still access these the same way
df[2,] # extracts the 2nd row, ALL columns
df[,2] # extracts ALL rows, the 2nd column
df$Name # This accesses the Name column

# However, you can do a lot more with Data Frames

df <- rbind(df, c("Rafael Nadal", NA)) # We can add a row
df <- cbind(df, Sport = c("Football", "Baseball", "Basketball", "Tennis")) # We can add a new column
df$Country <- c("USA", "USA", "USA", "Spain")  # We can reassign columns a lot easier
```

# Commenting Code

Commenting your code is the most important thing you can do. To comment in R, use a #. Commenting is super important because it allows you to make notes about functions you may be unfamiliar with or to explain your
reasoning to others. If you don't comment your code, you are a loser.


# Packages

```{r install_packages}
# install.packages('tidyverse') 
```
Packages must first be installed. Once you install it once you're good forever. That is why the previous line of code is commented out. To use a package, you need to use the *library()* function.

```{r library}
library(tidyverse) 
```

Tidyverse is the best overall collection of packages in R, includes ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr, & forcats. We will discuss some of these later.

# Introduction to Tidyverse

For more interesting data, we are going to use the Lahman datasets

```{r lahman}
#install.packages("Lahman")
library(Lahman)
```

Let's use some of the interesting functions in tidyverse

```{r select}
lahmanB <- Batting %>% 
  select(c(playerID, yearID, G:H, RBI, HR))
```
 %>% is called "pipe". It is a nicer way to nest functions and is available through the magrittr package which loads automatically when you library the tidyverse package. This code functions the same as:
 
```{r nopipe}
lahmanB2 <- select(Batting, c(playerID, G:R, RBI))
```

The select() function is from the dplyr package in tidyverse. It allows you to select certain variables for analysis.

```{r mutate}
lahmanP <- People %>% select(playerID, nameFirst, nameLast) %>% 
                      mutate(name = paste(nameFirst, nameLast))
```
The mutate function allows you to add columns to a data frame. The paste function allows you to concatenate strings together with a space between words.

```{r antiselect}
lahmanP <- lahmanP %>% select(-c(nameFirst, nameLast))
```
select can also be used to unselect columns as in the example above

```{r merge}
lahmanBig <- merge(lahmanB, lahmanP, by = "playerID")
```
The merge function takes two data sets and combines them by a similar column. This is similar to joining.

```{r reordercols}
lahmanBig <- lahmanBig %>% 
  select(name, playerID, everything())
```
The select function can also be used to reorder rows.The everything function gets all other variables in the same order. The everything function is restricted in when it can be used to selection functions, such as *select()*

```{r filter}
lahmanHR <- lahmanBig %>% 
              filter(HR >= 30) %>%
              arrange(desc(HR), RBI, name)
```
Filter limits data to only a certain characteristic. arrange sorts the data in the order of variables listed. desc() means descending. We sorted by descending number of HRs to see who hit the most in one season.

# R Documentation

* When you don't know how to do something, look at the R Documentation using a ? before the function you want to learn more about

```{r help}
# ?select 
```

# Conclusion

This wraps up the bare basics about coding in R. There are of courses many more things to learn both in other tutorials and through other resources. I want to say thanks to former NCSU Sports Analytics Club Board members Graham Pash (Twitter: @gtpash) and Jason Thompson (Twitter: @jason24thompson) for allowing me to build this tutorial series off of work they previously did.